---
title: "DMA with Sesame"
author: "Xiaoyu Yan"
date: "2/3/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sesame)
library(unpivotr)
library(tidyverse)
data_dir <- "~/Desktop/Graduate/MethData/"
```

# Tumor gland  

## Create sample directory file

```{r}
directory <- read_csv(paste0(data_dir,"directory.csv"))
directory <- directory %>% select(-type)

tumor_gland <- directory %>% filter(description == "tumor gland")
write.table(tumor_gland, file = "tumor_gland_dir.csv", quote = F, sep = ",", row.names = F)

# make tumor gland idat dir
tumor_gland_dir <- data.frame(union(paste("mkdir", paste0(data_dir, "tumor_gland")), 
                                    paste("cp", paste0(tumor_gland$sentrix_id, "_",
                                                       tumor_gland$sentrix_position, "*.*"), 
                                          paste0(data_dir, "tumor_gland"))
                                    )
                              )
write.table(tumor_gland_dir,file = "tumor_gland_dir.sh", quote = F, row.names = F, col.names = F)
```

## Idat file classification

```{bash}
cd ~/Downloads/MethData
sh ~/Desktop/Graduate/DMA/tumor_gland_dir.sh
```

## Read data & Get betas

```{r}
# Read IDATs into SigSet list
ssets_tumor_gland <- lapply(
  searchIDATprefixes(paste0(data_dir, "tumor_gland")), readIDATpair)
  
ssets <- sapply(
  ssets_tumor_gland, function(pfx) {
    pfx %>%
    pOOBAH %>% # sets the p-value for each probe
    noob #%>% # background subtraction based on normal-exponential deconvolution 
    #dyeBiasCorrTypeINorm # correction nonlinear quantile interpolation with Type-I probes
    }
  )

betas <- sapply(ssets, getBetas)
```

## Sample infer & QC

```{r}
# Sex infer
Sex <- vapply(ssets, inferSex, character(1))
SexKaryotypes <- vapply(ssets, inferSexKaryotypes, character(1))
# Ethnicity infer
Ethnicity <- vapply(ssets, inferEthnicity, character(1))
# Age infer
Age <- apply(betas, 2, predictAgeHorvath353)

# Mean intensity
meanIntensity <- vapply(ssets, meanIntensity, numeric(1))
# Bisulfite conversion control using GCT scores
# The closer the score to 1.0, the more complete the bisulfite conversion.
bcc <- vapply(ssets, bisConversionControl, numeric(1))
```

## Visualization

```{r}
# Visualize all probes from a gene
visualizeGene('DNMT1', betas, platform='EPIC')
# Visualize probes from arbitrary region
visualizeRegion(
  'chr19',10260000,10380000, betas, platform='EPIC',
  show.probeNames = FALSE)
# Visualize by probe names
visualizeProbes(c("cg02382400", "cg03738669"), betas, platform='EPIC')
# Copy number variation
# >50bp
# red -- del blu -- dup grn -- ins
ssets.normal <- sesameDataGet('EPIC.5.normal')
segs <- cnSegmentation(ssets[[1]], ssets.normal)
visualizeSegments(segs)

tumor_gland_ssets <- ssets
tumor_gland_betas <- betas
tumor_gland_segs <- segs
```
## Comparision

```{r}
A <- tumor_gland %>% filter(grepl("A", sample_id)) %>% select(plate, sentrix_id, sentrix_position)
B <- tumor_gland %>% filter(grepl("B", sample_id)) %>% select(plate, sentrix_id, sentrix_position) 
A
```

```{r}
CNV <- function(x) {cnSegmentation(ssets[[x]], ssets.normal) %>%
  visualizeSegments()
}
```

```{r}
CNV(1)
CNV(2)
```
```{r}
CNV(3)
CNV(4)
```

# Colon crypt

```{r}
colon_crypt <- directory %>%
  filter((grepl("colon", tissue)&grepl("crypt", description)))
write.table(colon_crypt, file = "colon_crypt_dir.csv", quote = F, sep = ",", row.names = F)

# make colon_crypt idat dir
colon_crypt_dir <- data.frame(union(paste("mkdir", paste0(data_dir, "colon_crypt")), 
                                    paste("cp", paste0(colon_crypt$sentrix_id, "_",
                                                       colon_crypt$sentrix_position, "*.*"), 
                                          paste0(data_dir, "colon_crypt"))
                                    )
                              )
write.table(colon_crypt_dir,file = "colon_crypt_dir.sh", quote = F, row.names = F, col.names = F)
```

## Idat file classification

```{bash}
cd ~/Downloads/MethData
sh ~/Desktop/Graduate/DMA/colon_crypt_dir.sh
```

## Read data & Get betas

```{r}
# Read IDATs into SigSet list
ssets_colon_crypt <- lapply(
  searchIDATprefixes(paste0(data_dir, "colon_crypt")), readIDATpair)
  
ssets <- sapply(
  ssets_colon_crypt, function(pfx) {
    pfx %>%
    pOOBAH %>% # sets the p-value for each probe
    noob #%>% # background subtraction based on normal-exponential deconvolution 
    #dyeBiasCorrTypeINorm # correction nonlinear quantile interpolation with Type-I probes
    }
  )

betas <- sapply(ssets, getBetas)
```

## Sample infer & QC

```{r}
# Sex infer
Sex <- vapply(ssets, inferSex, character(1))
SexKaryotypes <- vapply(ssets, inferSexKaryotypes, character(1))
# Ethnicity infer
Ethnicity <- vapply(ssets, inferEthnicity, character(1))
# Age infer
Age <- apply(betas, 2, predictAgeHorvath353)

# Mean intensity
meanIntensity <- vapply(ssets, meanIntensity, numeric(1))
# Bisulfite conversion control using GCT scores
# The closer the score to 1.0, the more complete the bisulfite conversion.
bcc <- vapply(ssets, bisConversionControl, numeric(1))
```

## Visualization

```{r}
# Visualize all probes from a gene
visualizeGene('DNMT1', betas, platform='EPIC')
# Visualize probes from arbitrary region
visualizeRegion(
  'chr19',10260000,10380000, betas, platform='EPIC',
  show.probeNames = FALSE)
# Visualize by probe names
visualizeProbes(c("cg02382400", "cg03738669"), betas, platform='EPIC')
# Copy number variation
# >50bp
# red -- del blu -- dup grn -- ins
ssets.normal <- sesameDataGet('EPIC.5.normal')
segs <- cnSegmentation(ssets[[1]], ssets.normal)
visualizeSegments(segs)
colon_crypt_ssets <- ssets
colon_crypt_betas <- betas
colon_crypt_segs <- segs
```


# SI crypt

```{r}
small_crypt <- directory %>% 
  filter(grepl("(SI crypt|crypt SI)", description))
write.table(small_crypt, file = "small_crypt_dir.csv", quote = F, sep = ",", row.names = F)

# make small_crypt idat dir
small_crypt_dir <- data.frame(union(paste("mkdir", paste0(data_dir, "small_crypt")), 
                                    paste("cp", paste0(small_crypt$sentrix_id, "_",
                                                       small_crypt$sentrix_position, "*.*"), 
                                          paste0(data_dir, "small_crypt"))
                                    )
                              )
write.table(small_crypt_dir,file = "small_crypt_dir.sh", quote = F, row.names = F, col.names = F)
```

## Idat file classification

```{bash}
cd ~/Downloads/MethData
sh ~/Desktop/Graduate/DMA/small_crypt_dir.sh
```

## Read data & Get betas

```{r}
# Read IDATs into SigSet list
ssets_small_crypt <- lapply(
  searchIDATprefixes(paste0(data_dir, "small_crypt")), readIDATpair)
  
ssets <- sapply(
  ssets_small_crypt, function(pfx) {
    pfx %>%
    pOOBAH %>% # sets the p-value for each probe
    noob #%>% # background subtraction based on normal-exponential deconvolution 
    #dyeBiasCorrTypeINorm # correction nonlinear quantile interpolation with Type-I probes
    }
  )

betas <- sapply(ssets, getBetas)
```

## Sample infer & QC

```{r}
# Sex infer
Sex <- vapply(ssets, inferSex, character(1))
SexKaryotypes <- vapply(ssets, inferSexKaryotypes, character(1))
# Ethnicity infer
Ethnicity <- vapply(ssets, inferEthnicity, character(1))
# Age infer
Age <- apply(betas, 2, predictAgeHorvath353)

# Mean intensity
meanIntensity <- vapply(ssets, meanIntensity, numeric(1))
# Bisulfite conversion control using GCT scores
# The closer the score to 1.0, the more complete the bisulfite conversion.
bcc <- vapply(ssets, bisConversionControl, numeric(1))
```

## Visualization

```{r}
# Visualize all probes from a gene
visualizeGene('DNMT1', betas, platform='EPIC')
# Visualize probes from arbitrary region
visualizeRegion(
  'chr19',10260000,10380000, betas, platform='EPIC',
  show.probeNames = FALSE)
# Visualize by probe names
visualizeProbes(c("cg02382400", "cg03738669"), betas, platform='EPIC')
# Copy number variation
# >50bp
# red -- del blu -- dup grn -- ins
ssets.normal <- sesameDataGet('EPIC.5.normal')
segs <- cnSegmentation(ssets[[1]], ssets.normal)
visualizeSegments(segs)
small_crypt_ssets <- ssets
small_crypt_betas <- betas
small_crypt_segs <- segs
```

# Endometrial Gland 

```{r}
endom_gland <- directory %>% 
  filter(grepl("(endo|endometrial)", description))
write.table(endom_gland, file = "endom_gland_dir.csv", quote = F, sep = ",", row.names = F)

# make endom_gland idat dir
endom_gland_dir <- data.frame(union(paste("mkdir", paste0(data_dir, "endom_gland")), 
                                    paste("cp", paste0(endom_gland$sentrix_id, "_",
                                                       endom_gland$sentrix_position, "*.*"), 
                                          paste0(data_dir, "endom_gland"))
                                    )
                              )
write.table(endom_gland_dir,file = "endom_gland_dir.sh", quote = F, row.names = F, col.names = F)
```

## Idat file classification

```{bash}
cd ~/Downloads/MethData
sh ~/Desktop/Graduate/DMA/endom_gland_dir.sh
```

## Read data & Get betas

```{r}
# Read IDATs into SigSet list
ssets_endom_gland <- lapply(
  searchIDATprefixes(paste0(data_dir, "endom_gland")), readIDATpair)
  
ssets <- sapply(
  ssets_endom_gland, function(pfx) {
    pfx %>%
    pOOBAH %>% # sets the p-value for each probe
    noob #%>% # background subtraction based on normal-exponential deconvolution 
    #dyeBiasCorrTypeINorm # correction nonlinear quantile interpolation with Type-I probes
    }
  )

betas <- sapply(ssets, getBetas)
```

## Sample infer & QC

```{r}
# Sex infer
Sex <- vapply(ssets, inferSex, character(1))
SexKaryotypes <- vapply(ssets, inferSexKaryotypes, character(1))
# Ethnicity infer
Ethnicity <- vapply(ssets, inferEthnicity, character(1))
# Age infer
Age <- apply(betas, 2, predictAgeHorvath353)

# Mean intensity
meanIntensity <- vapply(ssets, meanIntensity, numeric(1))
# Bisulfite conversion control using GCT scores
# The closer the score to 1.0, the more complete the bisulfite conversion.
bcc <- vapply(ssets, bisConversionControl, numeric(1))
```

## Visualization

```{r}
# Visualize all probes from a gene
visualizeGene('DNMT1', betas, platform='EPIC')
# Visualize probes from arbitrary region
visualizeRegion(
  'chr19',10260000,10380000, betas, platform='EPIC',
  show.probeNames = FALSE)
# Visualize by probe names
visualizeProbes(c("cg02382400", "cg03738669"), betas, platform='EPIC')
# Copy number variation
# >50bp
# red -- del blu -- dup grn -- ins
ssets.normal <- sesameDataGet('EPIC.5.normal')
segs <- cnSegmentation(ssets[[1]], ssets.normal)
visualizeSegments(segs)
endom_gland_ssets <- ssets
endom_gland_betas <- betas
endom_gland_segs <- segs
```

# Bulk Sample

```{r}
bulk_sample <- directory %>% filter(description == "bulk")
write.table(bulk_sample, file = "bulk_sample_dir.csv", quote = F, sep = ",", row.names = F)

# make bulk_sample idat dir
bulk_sample_dir <- data.frame(union(paste("mkdir", paste0(data_dir, "bulk_sample")), 
                                    paste("cp", paste0(bulk_sample$sentrix_id, "_",
                                                       bulk_sample$sentrix_position, "*.*"), 
                                          paste0(data_dir, "bulk_sample"))
                                    )
                              )
write.table(bulk_sample_dir,file = "bulk_sample_dir.sh", quote = F, row.names = F, col.names = F)
```

## Idat file classification

```{bash}
cd ~/Downloads/MethData
sh ~/Desktop/Graduate/DMA/bulk_sample_dir.sh
```

## Read data & Get betas

```{r}
# Read IDATs into SigSet list
ssets_bulk_sample <- lapply(
  searchIDATprefixes(paste0(data_dir, "bulk_sample")), readIDATpair)
  
ssets <- sapply(
  ssets_bulk_sample, function(pfx) {
    pfx %>%
    pOOBAH %>% # sets the p-value for each probe
    noob #%>% # background subtraction based on normal-exponential deconvolution 
    #dyeBiasCorrTypeINorm # correction nonlinear quantile interpolation with Type-I probes
    }
  )

betas <- sapply(ssets, getBetas)
```

## Sample infer & QC

```{r}
# Sex infer
Sex <- vapply(ssets, inferSex, character(1))
SexKaryotypes <- vapply(ssets, inferSexKaryotypes, character(1))
# Ethnicity infer
Ethnicity <- vapply(ssets, inferEthnicity, character(1))
# Age infer
Age <- apply(betas, 2, predictAgeHorvath353)

# Mean intensity
meanIntensity <- vapply(ssets, meanIntensity, numeric(1))
# Bisulfite conversion control using GCT scores
# The closer the score to 1.0, the more complete the bisulfite conversion.
bcc <- vapply(ssets, bisConversionControl, numeric(1))
```

## Visualization

```{r}
# Visualize all probes from a gene
visualizeGene('DNMT1', betas, platform='EPIC')
# Visualize probes from arbitrary region
visualizeRegion(
  'chr19',10260000,10380000, betas, platform='EPIC',
  show.probeNames = FALSE)
# Visualize by probe names
visualizeProbes(c("cg02382400", "cg03738669"), betas, platform='EPIC')
# Copy number variation
# >50bp
# red -- del blu -- dup grn -- ins
ssets.normal <- sesameDataGet('EPIC.5.normal')
segs <- cnSegmentation(ssets[[1]], ssets.normal)
visualizeSegments(segs)
bulk_sample_ssets <- ssets
bulk_sample_betas <- betas
bulk_sample_segs <- segs
```



